---
title: "Reproducible Research"
author: "tedding"
date: "April 12, 2016"
output: github_document
---




# Structure of a Data Analysis

##  steps in a data analysis
Define the question   
Define the ideal data set   
Determine what data you can access   
Obtain the data   
Clean the data   
Exploratory data analysis   
Statistical prediction/modeling   
Interpret results   
Challenge results   
Synthesize/ write up results   
Create reporducible code   


## What to Do and What Not to Do 
DO : Keep track of your software environment
```{r}
sessionInfo()
```

DON'T : only save the output, save data(data production flow) + code as well  
DO : set your seed, set.seed()  
DO : think about the entire pipeline    

- raw data -> processed data -> analysis -> report
- the more data analysis pipeline you make, the better for everyone


## Exploratory data analysis   
```{r, warning=FALSE}
library(kernlab)
data(spam)
set.seed(3435)
trainIndicator = rbinom( 4601, size = 1, prob = 0.5)
table(trainIndicator)
trainSpam = spam[ trainIndicator ==1, ]
testSpam = spam[ trainIndicator == 0, ]
# summarys, missing data, exploratory plots, clusting...
head(trainSpam)
plot( trainSpam$capitalAve ~ trainSpam$type )
plot( log10(trainSpam$capitalAve +1) ~ trainSpam$type ) # how to show data
plot( log10(trainSpam[,1:4]+1))
hCluster = hclust( dist(t(trainSpam[, 1:57])))
plot(hCluster)
# statistical prediction / modeling
trainSpam$numType = as.numeric( trainSpam$type) -1 
costFunction = function(x,y) sum( x!= (y>0.5))
cvError = rep(NA, 55)
library(boot)
for ( i in 1:55) {
  lmFormula = reformulate( names(trainSpam)[i], response = "numType")
  glmFit = glm( lmFormula, family = "binomial", data = trainSpam )
  cvError[i] = cv.glm( trainSpam, glmFit, costFunction, 2)$delta[2]
}

## which predictor has minimum cross-validated error?
names(trainSpam)[ which.min(cvError)]
## use the best model for the group
predictionModel = glm( numType ~ charDollar, family = "binomial", data = trainSpam )

## Get predictions on the test set
predictionTest = predict( predictionModel, testSpam)
predictedSpam = rep("nonspam", dim(testSpam)[1])

## classify as 'spam' for those with prob > 0.5
predictedSpam[ predictionModel$fitted > 0.5] = "spam"
##classification table
table( predictedSpam, testSpam$type)

```

# Organizing Your Analysis
Data: raw / precessed   
Figures:    
  <t>exploratory( do not need to be pretty)    
  <t>final( usually a small subse of original figs; axes,colors to make it clear)    
R code: raw / final / markdown   
test: read me / report   

# ref

- cheatsheet of rmarkdown

