---
title: "Practical Machine Learning week4"
author: "Dingchong"
date: "Monday, September 22, 2014"
output: html_document
---

# Week 4

## Regularized regression (13:20)
## Combining predictors (7:11)
## Forecasting
## Completed Unsupervised Prediction (4:24)


# Regularized regression

## Basic idea
Pros: help with bias/variance tradeoff; help with model selection   
Cons: computation demanding; does not perform as well as rf and boosting

## more vars, more overfit
```{r}

# regression subset selection in the prostate dataset
library(ElemStatLearn);data(prostate)
str(prostate)

covnames <- names(prostate[-(9:10)])
y <- prostate$lpsa
x <- prostate[,covnames]

form <- as.formula(paste("lpsa~", paste(covnames, collapse="+"), sep=""))
summary(lm(form, data=prostate[prostate$train,]))

set.seed(1)
train.ind <- sample(nrow(prostate), ceiling(nrow(prostate))/2)
y.test <- prostate$lpsa[-train.ind]
x.test <- x[-train.ind,]

y <- prostate$lpsa[train.ind]
x <- x[train.ind,]

p <- length(covnames)
rss <- list()
for (i in 1:p) {
  cat(i)
  Index <- combn(p,i)

  rss[[i]] <- apply(Index, 2, function(is) {
    form <- as.formula(paste("y~", paste(covnames[is], collapse="+"), sep=""))
    isfit <- lm(form, data=x)
    yhat <- predict(isfit)
    train.rss <- sum((y - yhat)^2)

    yhat <- predict(isfit, newdata=x.test)
    test.rss <- sum((y.test - yhat)^2)
    c(train.rss, test.rss)
  })
}

#png("Plots/selection-plots-01.png", height=432, width=432, pointsize=12)
plot(1:p, 1:p, type="n", ylim=range(unlist(rss)), xlim=c(0,p), xlab="number of predictors", ylab="residual sum of squares", main="Prostate cancer data")
for (i in 1:p) {
  points(rep(i-0.15, ncol(rss[[i]])), rss[[i]][1, ], col="blue")
  points(rep(i+0.15, ncol(rss[[i]])), rss[[i]][2, ], col="red")
}
minrss <- sapply(rss, function(x) min(x[1,]))
lines((1:p)-0.15, minrss, col="blue", lwd=1.7)
minrss <- sapply(rss, function(x) min(x[2,]))
lines((1:p)+0.15, minrss, col="red", lwd=1.7)
legend("topright", c("Train", "Test"), col=c("blue", "red"), pch=1)
#dev.off()
```
train error is smaller as new X adds, but test error is U shape.

## Model selection approach: split samples
### Approach 
1. Divide data into traing/test/validation    
2. Treat validation as test data, train all competing medels on the train data 
and pick the best one on validation.   
3. To appropriately assess performance on new data apply to test set   
4. you may re-split and reperform steps 1-3   
### this is similar as offline + online test

### Regularization for regression
1. Ridge regression- lamda   
2. Lasso
3. Relaxo

### Ridge regressioin
```{r}
# ridge regression on prostate dataset
library(MASS)
lambdas <- seq(0,50,len=10)
M <- length(lambdas)
train.rss <- rep(0,M)
test.rss <- rep(0,M)
betas <- matrix(0,ncol(x),M)
for(i in 1:M){
  Formula <-as.formula(paste("y~",paste(covnames,collapse="+"),sep=""))
  fit1 <- lm.ridge(Formula,data=x,lambda=lambdas[i])
  betas[,i] <- fit1$coef
  
  scaledX <- sweep(as.matrix(x),2,fit1$xm)
  scaledX <- sweep(scaledX,2,fit1$scale,"/")
  yhat <- scaledX%*%fit1$coef+fit1$ym
  train.rss[i] <- sum((y - yhat)^2)
  
  scaledX <- sweep(as.matrix(x.test),2,fit1$xm)
  scaledX <- sweep(scaledX,2,fit1$scale,"/")
  yhat <- scaledX%*%fit1$coef+fit1$ym
  test.rss[i] <- sum((y.test - yhat)^2)
}

#png(file="Plots/selection-plots-02.png", width=432, height=432, pointsize=12) 
plot(lambdas,test.rss,type="l",col="red",lwd=2,ylab="RSS",ylim=range(train.rss,test.rss))
lines(lambdas,train.rss,col="blue",lwd=2,lty=2)
best.lambda <- lambdas[which.min(test.rss)]
abline(v=best.lambda+1/9)
legend(30,30,c("Train","Test"),col=c("blue","red"),lty=c(2,1))
#dev.off()


#png(file="Plots/selection-plots-03.png", width=432, height=432, pointsize=8) 
plot(lambdas,betas[1,],ylim=range(betas),type="n",ylab="Coefficients")
for(i in 1:ncol(x))
  lines(lambdas,betas[i,],type="b",lty=i,pch=as.character(i))
abline(h=0)
legend("topright",covnames,pch=as.character(1:8))
#dev.off()
```

### lasso 
```{r}
# lasso
library(lars)
lasso.fit <- lars(as.matrix(x), y, type="lasso", trace=TRUE)
#png(file="Plots/selection-plots-04.png", width=432, height=432, pointsize=8) 
plot(lasso.fit, breaks=FALSE)
legend("topleft", covnames, pch=8, lty=1:length(covnames), col=1:length(covnames))
#dev.off()

# this plots the cross validation curve
#png(file="Plots/selection-plots-05.png", width=432, height=432, pointsize=12) 
lasso.cv <- cv.lars(as.matrix(x), y, K=10, type="lasso", trace=TRUE)
#dev.off()

````

# Combining predictors (7:11)
or ensembling
## Key ideas
1. combine classifiers by averaging / voting   
2. Boosting, bagging, and random forests are variants on this theme    
Netflix prize, BellKor = Combination of 107 predictors!!!   

## Approaches for combining classifiers
1. Bagging, boosting, random forests: combine similar classifiers   
2. Combining different classifiers: model stacking, model ensembling   
```{r}
library(ISLR); data(Wage); library(ggplot2); library( caret)
Wage <- subset(Wage, select=-c(logwage))
# create a building data set and validation set
inBuild <- createDataPartition( y= Wage$wage, p=0.7, list=F)
validation <- Wage[-inBuild, ]; buildData <- Wage[inBuild,]

inTrain <- createDataPartition( y=buildData$wage, p=0.7, list=F)
training <- buildData[inTrain, ]; testing <- buildData[-inTrain, ]

# Build 2 different models
mod1 <- train( wage ~., method="glm", data= training )
mod2 <- train( wage ~., method="rf", data= training, 
               trControl = trainControl( method="cv"), number=3 )

pred1 <- predict( mod1, testing); pred2 <- predict( mod2, testing)
qplot( pred1,pred2, color=wage, data=testing )

# combine predictors
predDf <- data.frame( pred1, pred2, wage=testing$wage)
combModFit <- train( wage ~., method="gam", data= predDf ) # 
combPred <- predict( combModFit, predDf)

```

## Notes
1. Build an odd number of models   
2. Predict with each model   
3. Predict the class by majority vote   
4. Scalability matters   


# Forecasting
Time series issues, ignored.

# Completed Unsupervised Prediction (4:24)
Ignored.

# Quiz 4

```{r}
## Q1, different predictors
library(ElemStatLearn);data(vowel.train);data(vowel.test) 
set.seed(33833)
head(vowel.test)
vowel.test$y <- as.factor(vowel.test$y)
vowel.train$y <- as.factor(vowel.train$y)
library(caret)
rf <- train( y ~., method ="rf", data = vowel.train)
gbm <- train( y~., method ="gbm", data = vowel.train )
rf.p <- predict( rf, vowel.test)
gbm.p <- predict( gbm, vowel.test)
confusionMatrix( rf.p, vowel.test$y )
confusionMatrix( gbm.p, vowel.test$y )

## Q2, combine models
library(caret)
library(gbm)
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
set.seed(62433)
rf <- train( diagnosis ~., method ="rf", data = training )
gbm <- train( diagnosis~., method ="gbm", data = training )
lda <- train( diagnosis~., method ="lda", data = training )
p.rf <- predict( rf, testing); p.gbm <- predict( gbm, testing); p.lda <- predict( lda, testing)
# combine predictors
predDf <- data.frame( p.rf, p.gbm, p.lda, y=testing$diagnosis)
combModFit <- train( y ~., method="rf", data= predDf ) # 
combPred <- predict( combModFit, predDf)
confusionMatrix( combPred, testing$diagnosis)
confusionMatrix( p.rf, testing$diagnosis)
confusionMatrix( p.gbm, testing$diagnosis)
confusionMatrix( p.lda, testing$diagnosis)

## Q3, lasso, weighting of vars
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
set.seed(233)
library(lars)
lasso.fit <- lars( as.matrix(training[,-9]), training$CompressiveStrength, type="lasso", trace=TRUE)
plot(lasso.fit, breaks=FALSE)
legend("topleft", names(training[,-9]), pch=8, lty=1:8, col=1:8)

## Q4
library(lubridate)  # For year() function below
dat <- read.csv("~/GitHub/datasciencecoursera/8.Practical Machine Learning/gaData.csv")
training = dat[year(dat$date) < 2012,]
testing = dat[(year(dat$date)) > 2011,]
tstrain = ts(training$visitsTumblr)

## Q5
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
set.seed(325)
library(e1071)
svm <- svm( CompressiveStrength~., data=training)
pred <- predict( svm, testing  )
sqrt( sum( (pred - testing$CompressiveStrength)^2 , na.rm = TRUE ) / nrow(testing) )

```

# References
## Statistical Learning: Algorithmic and Nonparametric Approaches
http://www.biostat.jhsph.edu/~ririzarr/Teaching/649/
