---
title: "[cousera???ݿ?ѧ]Exploratory Data Analysis "
author: "Dingchong"
date: "Friday, March 20, 2015"
output: github_document
---
## WEEK 1

### 1.Introduction


### 2.principles of analytic graphics
Principle 1: show comparisons

- Always ask "compared to what? "
- Related things

Principle 2: show causality, mechanism, explanation, systemiatic structure

- what is your causal framework for thinking about a question?

Principle 3: show multivariate data

- more than 2 variables
- ???Σ??㡢?ߡ????Σ?,??ͼ

Principle 4: Integration of evidence

- Completely integrate words, numbers, images, diagrams...

Principle 5: Describe with appropriate labels, scales, sources, etc


### 3.Exploratory Graphs
boxplot
histogram
scatterplot

### 4.Plotting Systems in R
3 core systems

1 the base plotting system

pros:
start with blank canvas and build up from there
start with plot function( or similar) 
use annotation functions to add/modify( text, lines, points, axis )

cons:
can't go back once plot has started( ie. to adjust margins); need to plan first
difficult to 'translate' to others once a new plot has been created( no graphical language)
plot is just a series of R commands

```{r}

library(datasets)
data(cars)
with( cars, plot(speed, dist))

```


2 the lattice system

pros:
single function call(xplot, bwplot)
most useful for conditiong types of plots: looking at how y changes with x across levels of z
automatically setting margins and spacing
good for many many plots on screen

cons:
cannot add to the plot once it is created

```{r}
library( lattice)
state = data.frame( state.x77, region = state.region)
xyplot( Life.Exp ~ Income | region , data = state , layout = c(4,1))

```

3 ggplot2 system

pros:
automatically deals with spacings, text...
easy to use than lattice
default is useful

```{r}
library(ggplot2)
data(mpg)
qplot( displ, hwy, data =mpg)

```

4 base plotting system

package:
1.graphics: plot, hist, boxplot...
2.grDevices: X11, PDF, PNG, PostScript...

do you need to dynamically resize the plot?

par() 

oma : outer margin size ( default is 0 for all sides )
```{r}
par("lty")
par("bg")
par("mar")
par("mfrow")

```

base ploting functions

plot
lines
points
text
title
mtext: margin text
aixs

```{r}
library(datasets)
with( airquality , plot( Wind, Ozone))
title( main = "Ozone and Wind in New York City ") 

# add points
with( subset(airquality , Month ==5), points( Wind, Ozone, col = "blue")) 
legend( "topright", pch =1, col = c("blue", "black"), 
        legend = c("May", "Other Months"))

model = lm(Ozone ~ Wind, airquality )
abline( model, lwd = 2)

```

multiple base plots

```{r}

par( mfrow = c(1,2), oma = c(0,0,2,0)) # oma outer margine

with( airquality, {
  plot( Wind, Ozone, main = "Ozone and Wind")
  plot( Solar.R, Ozone, main = "Ozone and Solar Radiation")
  mtext( "Ozone and Weather in New York City", outer = TRUE)
})


```


### 5.Graphics Devices in R





