---
[cousera数据科学]Exploratory Data Analysis 
---

## 第1课 向量化运算和debug

### 1.Introduction


### 2.principles of analytic graphics
Principle 1: show comparisons

- Always ask "compared to what? "
- Related things

Principle 2: show causality, mechanism, explanation, systemiatic structure

- what is your causal framework for thinking about a question?

Principle 3: show multivariate data

- more than 2 variables
- 多形（点、线、扇形）,多图

Principle 4: Integration of evidence

- Completely integrate words, numbers, images, diagrams...

Principle 5: Describe with appropriate labels, scales, sources, etc


### 3.Exploratory Graphs
boxplot
histogram
scatterplot

### 4.Plotting Systems in R
3 core systems

1 the base plotting system

pros:
start with blank canvas and build up from there
start with plot function( or similar) 
use annotation functions to add/modify( text, lines, points, axis )

cons:
can't go back once plot has started( ie. to adjust margins); need to plan first
difficult to 'translate' to others once a new plot has been created( no graphical language)
plot is just a series of R commands

```{r}

library(datasets)
data(cars)
with( cars, plot(speed, dist))

```


2 the lattice system

pros:
single function call(xplot, bwplot)
most useful for conditiong types of plots: looking at how y changes with x across levels of z
automatically setting margins and spacing
good for many many plots on screen

cons:
cannot add to the plot once it is created

```{r}
library( lattice)
state = data.frame( state.x77, region = state.region)
xyplot( Life.Exp ~ Income | region , data = state , layout = c(4,1))

```

3 ggplot2 system

pros:
automatically deals with spacings, text...
easy to use than lattice
default is useful

```{r}
library(ggplot2)
data(mpg)
qplot( displ, hwy, data =mpg)

```

4 base plotting system

package:
1.graphics: plot, hist, boxplot...
2.grDevices: X11, PDF, PNG, PostScript...

do you need to dynamically resize the plot?

par() 

oma : outer margin size ( default is 0 for all sides )
```{r}
par("lty")
par("bg")
par("mar")
par("mfrow")

```




