---
title: "exploratory data analysis 2"
author: "Dingchong"
date: "Tuesday, March 17, 2015"
output: github_document
---

## WEEK 2 Lattice and ggplot2

### 1.Latice Plotting System

xyplot: scatterplots, y~ x with condition argument

```{r}
library(lattice, datasets)

ariquality = transform( airquality , Month = factor(Month))
xyplot( Ozone ~ Wind | Month, data = ariquality, layout = c(5,1))

# panel functions


```


bwplot: box-and-whiskers plots( "boxlots")
historgram
stripplot : like boxplot but with actual points
dotplot : plot dots on 'violin strings'
splom : scatterplot matrix: like pairs in base plotting system
levelplot, contourplot : "image" data


### 2.ggplot2 

basice: qplot()
```{r}
library(ggplot2)
str(mpg)

# 2 vaiable : scatter
qplot( displ, hwy, data = mpg, color = drv )

qplot( displ, hwy, data = mpg, geom = c("point", "smooth"))

qplot( displ, hwy, data = mpg, geom = c("point", "smooth"), method = "lm")


# 1 variable : bar
qplot( hwy, data = mpg, fill = drv )
qplot( log(hwy) , data = mpg, fill = drv )
qplot( log(hwy) , data = mpg, fill = drv, geom = "density" )
qplot( log(hwy) , data = mpg, color = drv, geom = "density" )

# subsets, with facets
qplot( displ, hwy, data = mpg, facets = .~drv )

qplot( hwy, data = mpg, facets = .~drv, binwidth =2  ) 
qplot( hwy, data = mpg, facets = drv~. , binwidth =2  )



```

basic components of a ggplot2 

A data frame
aesthetic mappings : how data are mapped to color, size 
geoms : geometric objects like points, lines, shapes.
facets: for conditional plots 
stats : statistical transformations like binning, quantiles, smoothing. 
scales : what scale an aesthetic map uses ( example : male = red, female = blue ) 
coordinate system 


A notes about axis limits


```{r}

testdat = data.frame( x=1:100, y = rnorm(100))
testdat[50,2] = 100 ## outlier
plot( testdat$x, testdat$y, type ="l", ylim = c(-3,3))


g = ggplot( testdat, aes(x =x, y=y))
g + geom_line()

# outlier missing
g = ggplot( testdat, aes(x =x, y=y))
g + geom_line() + ylim( -3, 3)
# outlier included
g = ggplot( testdat, aes(x =x, y=y))
g + geom_line() + coord_cartesian( ylim = c(-3,3))

```

a full story

```{r}

str(mpg)
g = ggplot( mpg, aes(x =displ, y= hwy)) 
g + geom_point( alpha = 1/3 )  + # add points
  facet_wrap( drv ~ cyl) + # factors in different color
  geom_smooth( method ="lm", col = "steelblue ") + # smoother
  theme_bw( base_family = "", base_size = 20 ) + # size, character
  labs( x = expression( "log" )) + 
  labs( y = "Nocturnal Symptons") + 
  labs( title = "MAACS Cohort" )


```








